---
title: "Lab 2 Homework Template"
author: "Your Name Here"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # don't show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = "center"   # center figures
)

```


```{r include=FALSE}
library(behaviouR)
library(ggpubr)

```

## Part 1. Enter and visualize your ethogram data

**Question 1**. Change the code below to reflect the categories you used in your ethogram and change the TimesObserved values to the actual values you recorded. NOTE: You may need to add more categories if your ethogram had more than six. What trends did you notice in terms of behaviors observed?


```{r echo=TRUE}

# MODIFY THE CODE BELOW
EthogramDFupdated <- data.frame(Behavior=c('Category1','Category2','Category3',
                                           'Category4','Category5','Category6'),
                         TimesObserved=c(1,2,3,4,5,6))

# Now plot the results
ggbarplot(data=EthogramDFupdated, x='Behavior', y='TimesObserved',fill='grey')
```


## Part 2. Calculate meerkat activity budgets. 
Now we compare our data with our partner's. Assuming you were able to successfully upload your data the code below should run without problems.


```{r echo = FALSE}
data("MeerkatFocalData")
MeerkatFocalDataPartner <- MeerkatFocalData %>% 
  dplyr::group_by(BehaviorCode,Partner) %>% 
  dplyr::summarise(TotalSeconds = sum(SecondsEngagedinBehavior))


#Again, we divide by 600 because our video was 10 minutes (or 600 seconds) long. # We then multiply by 100 so that we can report in percentages.

MeerkatFocalDataPartner$ActivityBudget <- MeerkatFocalDataPartner$TotalSeconds/600*100


# Note that there are two new lines in the code below. The fill argument tells R which category or factor to use to color the bars. The position argument tells R to place the bars side-by-side.
ggbarplot(data=MeerkatFocalDataPartner,x='BehaviorCode',y='ActivityBudget', fill='Partner',
          position = position_dodge(0.9))+
  ylab('Percent of Time')
```
  
  
**Question 2**. Were there any noticeable differences between you and your partner's activity budgets?  

ADD TEXT HERE TO ANSWER QUESTION 2
  
  
## Part 3. Scan sampling and inter-observer reliability.
Now we will compare inter-observer reliability. Assuming you were able to successfully upload your data the code below should run without problems.


```{r tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE}
# Read in the data
data("MeerkatScanData")

# Remove the first column
MeerkatScanData <-MeerkatScanData[,-c(1)]

# Convert zeros to NAs
MeerkatScanData[is.na(MeerkatScanData)] <- 0

# Now we summarize by partner
MeerkatScanDataSummary <- MeerkatScanData %>% 
  dplyr::group_by(Treatment,Partner) %>% 
  dplyr::summarise(Vigilant = sum(Vigilant),
                   OutOfSight=sum(OutOfSight),
                   NotVigilant=sum(NotVigilant))

#There are many functions in R that allow us to change the format of our data. This one changes the format to one that is more easily used by 'ggpubr'
MeerkatScanDataSummaryLong <- reshape2::melt(MeerkatScanDataSummary, id.vars=c("Treatment", "Partner"))

#Now we change the column names so they are better for plotting
colnames(MeerkatScanDataSummaryLong) <- c('Treatment','Partner','Behavior','NumberObservations')

# And now we can visually compare our results with our partners
ggbarplot(data=MeerkatScanDataSummaryLong,x='Treatment',y='NumberObservations', fill='Behavior',
          position = position_dodge(0.9),facet.by = 'Partner')
```

Now we will calculate inter-observer reliability. You may need to modify the code below to reflect the names of you and your partner. 

```{r echo = FALSE}
Partner1Data <- subset(MeerkatScanData,Partner=='A')
Partner1Data <- Partner1Data[,-c(5)] # We remove the 'Partner' column so that we only focus on scans

Partner2Data <- subset(MeerkatScanData,Partner=='B')
Partner2Data <- Partner2Data[,-c(5)] # We remove the 'Partner' column so that we only focus on scans

# Here we calculate reliability for vigilance
VigilantCorrelation <- cor(Partner1Data$Vigilant,Partner2Data$Vigilant)

# Here we calculate reliability for vigilance
NotVigilantCorrelation <- cor(Partner1Data$NotVigilant,Partner2Data$NotVigilant)

# Here we calculate reliability for out of sight
OutOfSightCorrelation <-  cor(Partner1Data$OutOfSight,Partner2Data$OutOfSight)

# Print the results
cbind.data.frame(VigilantCorrelation,NotVigilantCorrelation,OutOfSightCorrelation)


```
  
  
**Question 3**. What was the reliability (or correlation coefficent) between you and your partner for each of the different behavioral categories? What do you think lead to these differences?    


ADD TEXT HERE TO ANSWER QUESTION 3
