---
title: "Lab 4 Homework Template"
author: "Your Name Here"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
---

```{r include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # don't show code
  warning = FALSE,       # don't show warnings
  message = FALSE,       # don't show messages (less serious warnings)
  cache = FALSE,         # set to TRUE to save results from last compilation
  fig.align = "center"   # center figures
)

```

```{r include = FALSE}
# Load the libraries
library(behaviouR)
library(ggpubr)
library(lme4)
library(bbmle)
```


### Part 1a: Surveillance behavior
**Question 1**: Were there any major differences between you and your partner in terms of the observed relationship between flock size and surveillance rate?  

YOUR ANSWER HERE

**Question 2**: If the null model is ranked higher than the model with flock size as a predictor, how do would we interpret this finding? What if the model with flock size as a predictor was ranked higher? What were your results?  

YOUR ANSWER HERE

### Part 1b: Time vigilant (sec/min)
NOTE: You should not have to modify the code below to run assuming your barnacle goose data is in the correct format!

```{r echo=TRUE}
BarnacleGooseData <- read.csv('BarnacleGooseData.csv')

ggscatter(data=BarnacleGooseData,x='FlockSize',y='TimeSecHeadUp',add='reg.line',
          facet.by = 'Partner', color='Partner', cor.coef = T,
          palette =c('black','blue'))+
            ylab('time vigilant (sec/min)')

# This is our null model
VigilanceNullModel <- lmer(TimeSecHeadUp ~ (1|Partner), data=BarnacleGooseData)

# This is our model with flock size as a predictor duration of vigilance
VigilanceModel <- lmer(TimeSecHeadUp ~ FlockSize + (1|Partner) ,data=BarnacleGooseData)

# Now we compare the models using AIC
AICtab(VigilanceNullModel,VigilanceModel)
```

**Question 3.** How do you interpret the results of your model selection? Was there a relationship between flock size and duration of vigilance behavior?  

YOUR ANSWER HERE

### Part 2: Meerkat data revisted

**Question 4.** Based on your inspection of the boxplots, are there any major differences between treatment groups?  

YOUR ANSWER HERE   


```{r echo=FALSE}
# Read in data
MeerkatScanData <- read.csv('MeerkatScanData.csv')

# As before we will turn our NA values to zero
MeerkatScanData[is.na(MeerkatScanData)] <- '0'

# We will remove the time and out of sight columns as we do not need them
MeerkatScanData <- dplyr::select(MeerkatScanData,-c(Time,OutOfSight))

# We need to reformat our data so that we can plot it
MeerkatScanDataSummaryLong <- reshape2::melt(MeerkatScanData, id.vars=c("Treatment", "Partner"))

# Here we add more informative column names
colnames(MeerkatScanDataSummaryLong) <- c('Treatment','Partner',
                                          'BehavioralState','InstancesOfBehavior')

# We need to tell R that our outcome variable is not categorical but numeric
MeerkatScanDataSummaryLong$InstancesOfBehavior <- 
  as.numeric(MeerkatScanDataSummaryLong$InstancesOfBehavior)

# First we subset our data so that it only includes the vigilant category
MeerkatScanDataVigilantOnly <- subset(MeerkatScanDataSummaryLong,
                                      BehavioralState=='Vigilant' )

# R can be picky about the format of data, so we use this command to tell R that treatment group is a factor
MeerkatScanDataVigilantOnly$Treatment <- 
  as.factor(MeerkatScanDataVigilantOnly$Treatment)

# Here we are reordering the levels of the factors. For our model selection we are interested
# in whether we see differences from the control (no predator) and the predator treatments,
# so here we are setting the no predator group as our reference group.
MeerkatScanDataVigilantOnly$Treatment <- 
  factor(MeerkatScanDataVigilantOnly$Treatment, levels = c("NoPredator", "AerialPredator", 
                                                           "TerrestrialPredator"))

# Now as before we will do model selection. Note that because our outcome variable (instances of behavior) 
# is in the form of count data we use a poisson distribution.
MeerkatVigilanceNullModel <- glm(InstancesOfBehavior ~ 1, family=poisson, data=MeerkatScanDataVigilantOnly)
MeerkatVigilanceModel <- glm(InstancesOfBehavior ~ Treatment, family=poisson,data=MeerkatScanDataVigilantOnly)

# Now we compare the models using AIC
AICtab(MeerkatVigilanceNullModel,MeerkatVigilanceModel)

coefplot::coefplot(MeerkatVigilanceModel,intercept=FALSE)
```

**Question 5.** Based on your interpretation of the model selection and the coefficient plots were there differences between treatment groups (e.g. control, terrestrial and aerial predators) in meerkat vigilance behavior?  

YOUR ANSWER HERE
